Gravitational field calculation in spherical coordinates using variable densities in depth

### 2018-09-20 manuscript/manuscript.tex
Start editing algorithm section

We present a new algorithm for computing the gravitational fields generated by any
tesseroid with an arbitrary continuous density function on any external point.
It is based on the GLQ approximation and the adaptive discretization of
\citep{Uieda2016}, which we extend to include a new density-based discretization step.
In order to ensure the accuracy of the numerical approximation we have
determined optimal values for the controlling parameters by
comparing the numerical approximation with analytical solutions for
spherical shells.
Finally, we applied the methodology to model the Nequen basin, Argentina, using
tesseroids with an exponentially increasing density with depth.

@@

\noindent $(r_i, \phi_j, \lambda_k)$ are the GLQ nodes, and $W_i^r$ are
the quadrature weights.

@@

Studying the homogeneous density tesseroid problem, \citet{Ku1977}
noticed that keeping the GLQ orders fixed, the approximation
becomes less accurate when the computation point is closer to the
tesseroid, making the point masses effect more noticeable than the
tesseroid we want to approximate.
One way to prevent this could be increasing the GLQ order in case that
the computation point is near the tesseroid.
This would uniformly increase the number of point masses inside the
tesseroid volume.

Instead of doing this, \citet{Li2011} proposed an adaptive
discretization algorithm which essentially consists in dividing the
tesseroid based on a ratio between the distance to the computation
point and its dimensions.
This algorithm produces a more efficient computation because the
increase of point masses is produced where are more needed.

\citet{Uieda2016} proposed a modified version of this algorithm, whose
modifications can be summed up in:
(1) a faster calculation of the distance from the computation point to
the tesseroid and
(2) the application of a stack based algorithm that speeds up the
computation and gives more control over the recursion step.
More specifically, \citet{Uieda2016} redefined the dimensions of an
arbitrary tesseroid: $L_\lambda$, $L_\phi$ and $L_r$. The former ones
are the arc-distances measured along its top surface while $L_r$ is the
difference between the radii of the top and bottom surfaces.

>>>>

We present a new algorithm for computing the gravitational fields generated by any
tesseroid with an arbitrary continuous density function on any external point.
It is based on the GLQ approximation and the adaptive discretization of
\citet{Uieda2016}, which we extend to include a new density-based discretization step.
In order to ensure the accuracy of the numerical approximation we have
determined optimal values for the controlling parameters by
comparing the numerical approximation with analytical solutions for
spherical shells.
Finally, we applied the methodology to model the Nequen basin, Argentina, using
tesseroids with an exponentially increasing density with depth.

@@

\noindent $(r_i, \phi_j, \lambda_k)$ are the quadrature nodes, and $W_i^r$ are the
quadrature weights.

@@

\citet{Ku1977} noticed that the GLQ integration
becomes less accurate when the computation point is closer to the
mass element.
One way to prevent this from happening would be to increase the GLQ order.
Doing so would uniformly increase the number of point masses inside the
tesseroid volume.
However, an increase in the point mass concentration is only required close to the
computation point \citep{Uieda2016}.
Alternatively, \citet{Li2011} proposed an adaptive
discretization algorithm which keeps the GLQ order fixed and divides the
tesseroid based on a ratio between the distance to the computation
point and its dimensions.
This algorithm produces a more efficient computation because an increased concentration
of point masses is produced where it is needed more.
\citet{Uieda2016} developed a modified version of this algorithm, which we will use
here.
What follows is a summary of the algorithm and the reader is referred to
\citet{Uieda2016} for a detailed description.

The following algorithm computes the gravitational fields at a given point at $(r, \phi,
\lambda)$ due to a tesseroid with its geometric center at $(r_t, \phi_t, \lambda_t)$ and
dimensions $L_r$, $L_\phi$, and $L_\lambda$ given by

\begin{equation}
    L_\lambda = r_2 \arccos(\sin^2\phi_t +
        \cos^2\phi_t\cos(\lambda_2 - \lambda_1)),
    \label{eq:sizelon}
\end{equation}

\begin{equation}
    L_\phi = r_2 \arccos(\sin\phi_2\sin\phi_1 + \cos\phi_2\cos\phi_1),
\end{equation}

and

\begin{equation}
    L_r = r_2 - r_1.
    \label{eq:sizer}
\end{equation}

\noindent
\textit{Step 1}: Check that the tesseroid satisfies the following inequality for each
dimension $L_i$ of the tesseroid:

\begin{equation}
    \frac{d}{L_i} \geq D,
    \label{eq:condition}
\end{equation}

\noindent
in which $D$ is a positive scalar called the distance-size ratio and $d$ is the distance
between the computation point and the geometric center of the tesseroid

\begin{equation}
    d = \left[
        r^2 + r_t^2 - 2 r r_t \cos\psi_t
        \right]^{\frac{1}{2}} ,
    \label{eq:distance}
\end{equation}

\begin{equation}
    \cos\psi_t =
        \sin\phi\sin\phi_t + \cos\phi\cos\phi_t\cos(\lambda - \lambda_t) .
\end{equation}

\noindent
\textit{Step 2}: If none of the dimensions of the tesseroid fail inequality
\ref{eq:condition}, then compute the gravitational effect of the tesseroid using a
second-order GLQ (Eq. \ref{eq:glq-var-dens})
